---
title: "Project Documentation"
output: html_notebook
---
Feb 14th
```{r}
parse_date("4/1/2015", "%m/%d/%Y")
# parses our date successfully

parse_time("0:21:00")
```
```{r}
# This is for future reference, for now we are good.
uber$date <- as.Date(data$Date/Time) # add date column
uber$time <- format(as.POSIXct(data$Date/Time),format = "%H:%M:%S") # add time column
# source: https://statisticsglobe.com/split-date-time-column-into-separate-variables-in-r

```

Feb 16th
```{r}
# separating date and time from Date/Time column

new_uber <- uber %>%
  separate('Date/Time', into = c("date", "time"), sep = " ")
```

Feb 17th
Ran into this error when playing around: Discrete value supplied to continuous scale. This happens when applying aes to a categorical value when I put the column name in quotes.

Also ran into a problem where get_map() requires additional API authorization but it is resolved by lifting the restriction from the cloud platform.

```{r}
ggmap(nyc_map) + 
    geom_point(data = uber, aes(x= Lon,y= Lat), size=0.15,alpha=.3)
```
```{r}
#Run this line before using ggmap
register_google(key = "AIzaSyAnGi82u7r4pOg8fYXC_dugPcMeiA2OMf8")
```

```{r}
nyc.map <- get_map(location= 'Lower Manhattan, New York', 
                       maptype='roadmap', color='color',source='google',zoom=14)

ggmap(nyc.map) + 
    geom_point(data = uber, aes(x= Lon,y= Lat), size=0.15,alpha=.3)
```

Feb 18th
```{r}
# changed the date format to accomendate sliderInput in Shiny
new_uber <- uber %>%
  separate('Date/Time', into = c("date", "time"), sep = " ") %>%
  mutate(date = mdy(date))
```

I was looking into how to connect the date range input to the dataset, but was having a little trouble doing that, might need more research. 
```{r}
server <- function(input, output) {

    output$pickupPlot <- renderPlot({
        nyc.map <- get_map(location= 'Lower Manhattan, New York', 
                           maptype='roadmap', color='color',source='google', zoom = input$zoom)
        
        ggmap(nyc.map) + 
            geom_point(data = uber, aes(x= Lon,y= Lat), size=0.3,alpha=.3)
        # specifically select 'uber' based on the date from the input.
        
    })
}
```



